\textbf{COMP0086 Summative Assignment}

### 2022-11-11 submission.tex
test overleaf
Updates from Overleaf
Main

\documentclass[12pt]{article}
\usepackage[paper=letterpaper,margin=2cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{newtxtext, newtxmath}
\usepackage{enumitem}
\usepackage{titling}
\usepackage[colorlinks=true]{hyperref}
\usepackage{graphicx}
\usepackage{csvsimple}

\setlength{\droptitle}{-6em}

@@

\title{\textbf{COMP0086 Summative Assignment}}
\author{Alexei Pisacane \& Jian Shu (James) Wu \\ }
\date{Nov 16, 2022}

@@

        \item Plotting the $\ln$ of the test error for a single trial of 1000 test points versus the polynomial dimension $k = 1, . . . , 18$:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q2/q2c}
            \caption{Testing Error of Polynomial Curves (1 trial)}
            \label{fig:2c}
            \end{figure}
\newpage
        \item Plotting the $\ln$ of the test error for 100 trials of 1000 test points versus the polynomial dimension $k = 1, . . . , 18$:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q2/q2d}
            \caption{Testing Error of Polynomial Curves (100 trials)}
            \label{fig:2d}
            \end{figure}
    \end{enumerate}
\newpage
\item[3.] Repeating 2 (b-d) with $\sin$ basis functions:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q3/q3b}
            \caption{Training Error of $\sin$ Curves}
            \label{fig:3b}
            \end{figure}
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q3/q3c}
            \caption{Testing Error of $\sin$ Curves (1 trial)}
            \label{fig:3c}
            \end{figure}
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q3/q3d}
            \caption{Testing Error of $\sin$ Curves (100 trials)}
            \label{fig:3d}
            \end{figure}
\newpage
\subsection{Filtered Boston housing and kernels}
\item[4.]
\newpage
\subsection{Kernelised ridge regression}

@@

\newpage
\section{PART II}
\subsection{$k$-Nearest Neighbors}
\item[6.]Generating a random $h$:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q6/q6}
            \caption{A hypothesis $h_{S,v}$ visualised with $|S|=100$ and $v=3$}
            \label{fig:6a}
            \end{figure}
\newpage
\item[7.]
    \begin{enumerate}
        \item Producing a visualisation using Protocol A:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q7/q7}
            \caption{Protocol A}
            \label{fig:7}
            \end{figure}
        \item We know that the data generating process uses $v=3$, so we would have initially expected the smallest test error to be located at the matching k-value, $k=3$.
        However, the data generating process for the training data set labels \textit{and} the testing data set labels were injected with noise, so k-values greater than 3 have a ``de-noising'' effect allowing for better performance at around $k=7$.
        As the k-value increases beyond $k=7$, this smoothing effect is too strong, causing the gradual increase in test error as the k value increases.
        For $k<7$ and in particular $k<v=3$ we see the error growing because the model has small $k$ and is able to fit to the noise injected into the data generating process, corrupting the model.
        By averaging over 100 runs, the plot is relatively smooth allowing us to make generalisations about the relationship between k-value and mean test set error.
    \end{enumerate}
\newpage
\item[8.]
    \begin{enumerate}
        \item Producing a visualisation using Protocol B:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q8/q8}
            \caption{Protocol B}
            \label{fig:8}
            \end{figure}
        \item k-nearest neighbours is a non-parametric model, so when queried with a test point for prediction, the number of training points within a given fixed radius of that test point increases as the number of training points increases.
        If we consider this fixed radius `sufficiently close' to make an informed prediction, the number of points within this radius will increase with more training points.
        So as the number of training points increases, the optimal k would increase to include more of the training points within this same fixed radius.
        We can see in the plot above that we see this general trend as expected, however there is considerable noise in the trend and so a smoother plot would likely requires more runs.
    \end{enumerate}
\newpage
\section{PART III}
\subsection{Questions}
\item[9.] \begin{enumerate}
              \item let $K_{c}(\textbf{x},\textbf{z}) = c+ \sum_{i=1}^{n}x_{i}z_{i} = c + \textbf{x}^{T}\textbf{z}$\\
        Our function $K_{c}(\cdot, \cdot)$ is psd iff:\\
        $\forall \{ u_{i} \}_{1:m} \in \mathcal{R},
         \{ \textbf{x}_{j}\}_{1:m} \in \mathcal{R}^{n}$:

        $\sum_{k,l} u_{k}u_{l}K_{c}(\textbf{x}_{u},\textbf{x}_{l}) \ge 0$
        We expand our kernel function to give: \\

        \[\sum_{k,l} u_{k}u_{l}K_{c}(\textbf{x}_{u},\textbf{x}_{l}) =  \sum_{n,m} cu_{n} u_{m} + \sum_{n,m} \langle\textbf{x}_{n}, \textbf{x}_{m}\rangle u_{n} u_{m}\]

@@

        \textbf{Proposition}

        $K_{c}(\cdot, \cdot)$ is psd iff $c \ge0$

        \textbf{Proof:}\\
        $(\implies)$\\

@@

    We note that as c becomes large, $\alpha \to 0$ and hence the predictions for $\hat{y}_{test}$ approach
    the constant function $f(x) = lc$
\end{enumerate}

@@

    $K = I_{l} + \tilde{K}$, where $\tilde{K}_{ij} = K(x_i, x_j)$, $i \ne j$. We note here that as $\beta \to \infty$, $ \tilde{K} \to 0$.\\

@@

    Let's assume that there exists a unique point x in our training set of minimal distance to our test point.\\

    Define $d_{i} := \|x_{test} - x_{i}\|^{2}$, and $d* := min_{i} d_{i}$\\
    \\
    $g(t) = \frac{y_{*} exp(-\beta d_{*})}{|\sum_{i} y_{i} exp(-\beta d_{i})|} + \sum_{i\ne *}\frac{y_{i} exp(-\beta d_{i})}{|\sum_{i} y_{i} exp(-\beta 
    d_{i})|} = \frac{y_{*}}{|y_{*} + \sum_{i \ne *} y_{i} exp(-\beta (d_{i} - d_{*}))|} + \sum_{i\ne *}\frac{y_{i} exp(-\beta (d_{i} - d_{*}))}{|y_{*} + \sum_{i \ne *} y_{i} exp(-\beta (d_{i} - d_{*}))|}$.\\

    Note for $d_{i} \ne d_{*}$, $exp(-\beta(d_{i} - d_{*})) \to 0$ as $\beta \to \infty$. Hence:\\
    $g(t) \to \frac{y_{*}}{|y_{*} + 0|} + \frac{0}{|y_{*} + 0|} = y_{*} = y_{i} : x_{i} = min_{i} \|x_{test} - x_{i}\|$ which is the required 1-NN predictor.\\$
    \square$
\newpage
\item[11.] Given a whack-a-mole grid of size n, it is solvable if we can apply a sequence from our $n^{2}$ permissible moves that results in a blank screen. We first index all permissible moves by $T_{i,j}$, which has the effect of flipping the state of squares $(i,j),(i-1,j),(i+1,j),(i,j-1)(i,j+1)$, as long as each coordinate is still on the grid.\\
    We may consider each grid as an element of the space of square matrices over $\mathcal{F}_{2}$, the finite field of order 2: under this construction, we see that the application of a move $T_{i,j}$ corresponds to the sum between the grid and a matrix $T_{ij}$, where $T_{ij}$, where the entries of $T_{ij}$ are zero everywhere except at $(i,j),(i-1,j),(i+1,j),(i,j-1)(i,j+1)$, if these values are present on the grid. For example, if n = 4, the transformation $T_{2,2}$ corresponds to the addition of the matrix $T_{22}$:\\

>>>>

\documentclass[12pt]{article}
\usepackage[paper=letterpaper,margin=2cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{newtxtext, newtxmath}
\usepackage{enumitem}
\usepackage{titling}
\usepackage[colorlinks=true]{hyperref}
\usepackage{graphicx}
\usepackage{csvsimple}
\usepackage{luacode}
\setlength{\droptitle}{-6em}

@@

\title{\textbf{COMP0086 Summative Assignment}}
\author{Alexei Pisacane and Jian Shu (James) Wu \\ }
\date{Nov 16, 2022}

@@

        \item Plotting the $\ln$ of the test error for a single trial of 1000 test points versus the polynomial dimension $k = 1, . . . , 18$:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q2/q2c}
            \caption{Testing Error of Polynomial Curves (1 trial)}
            \label{fig:2c}
            \end{figure}
\newpage
        \item Plotting the $\ln$ of the test error for 100 trials of 1000 test points versus the polynomial dimension $k = 1, . . . , 18$:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q2/q2d}
            \caption{Testing Error of Polynomial Curves (100 trials)}
            \label{fig:2d}
            \end{figure}
    \end{enumerate}
\newpage
\item[3.] Repeating 2 (b-d) with $\sin$ basis functions:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q3/q3b}
            \caption{Training Error of $\sin$ Curves}
            \label{fig:3b}
            \end{figure}
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q3/q3c}
            \caption{Testing Error of $\sin$ Curves (1 trial)}
            \label{fig:3c}
            \end{figure}
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q3/q3d}
            \caption{Testing Error of $\sin$ Curves (100 trials)}
            \label{fig:3d}
            \end{figure}
\newpage
\subsection{Filtered Boston housing and kernels}
    \item[4.]
    \begin{enumerate}
        \item 
        We interpret the naive linear regression model using the existing framework for least squares regressors.\\
        Namely, we have:\\
        $w^* = (X^TX)^{-1}X^{T} y, X = (1,\dots,1)^{T} $\\
        \[\implies w* = m^{-1}  \sum_{i} 1*y_{i} = \bar{y}\]\\
        \[\implies f(x_{i}) = 1 \times w^{*} = \bar{y}\]\\
        We split the data using our train test split function, and train naive, single and multiple linear regression models.
            We repeat this 20 times and report the MSE on training and test sets.\\

        Here are the plots of training and test error:\\
        \includegraphics[scale=0.5]{outputs/q4/q4_train.png}\\
        \includegraphics[scale=0.5]{outputs/q4/q4_test.png}\\
    \end{enumerate}
\newpage
\subsection{Kernelised ridge regression}
\item[5.]
We fit our kernel ridge regressors for each parameter pair over 5 cross validation folds,
and search for the pair minimising our mean squared error.\\

Here are the optimal parameters along with the minimal mean squared error:\\

\item The mean squared error for each fitted curve:
\begin{center}
\begin{tabular}{c|c|c}%
\textbf{Gamma}&\textbf{Sigma}&\textbf{MSE}% specify table head
\csvreader[head to column names]{outputs/q5/optimal_gkrr_parameters.csv}{}% use head of csv as column names
{\\\hline\csvcoli&\csvcolii&\csvcoliii}% specify your coloumns here
\end{tabular}
\end{center}

\\
Here is our contour plot for mse over our parameter grid:\\

\includegraphics[scale = 0.5]{outputs/q5/contour_plot.png}\\

@@

Here is our final regression report over all oour models, run 20 times each.\\

\item The mean squared error for each fitted curve:
\begin{center}
\begin{tabular}{c|c|c|c|c}%
\textbd{model}&\textbf{train mse}&\textbf{train std}&\textbf{test mse}&\textbf{test std}% specify table head
\csvreader[head to column names]{outputs/q5/regression_report.csv}{}% use head of csv as column names
{\\\hline\csvcoli&\csvcolii&\csvcoliii&\csvcoliv&\csvcolv}% specify your coloumns here
\end{tabular}
\end{center}

\section{PART II}
\subsection{$k$-Nearest Neighbors}
\item[6.]Generating a random $h$:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q6/q6}
            \caption{A hypothesis $h_{S,v}$ visualised with $|S|=100$ and $v=3$}
            \label{fig:6a}
            \end{figure}
\newpage
\item[7.]
    \begin{enumerate}
        \item Producing a visualisation using Protocol A:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q7/q7}
            \caption{Protocol A}
            \label{fig:7}
            \end{figure}
        \item We know that the data generating process uses $v=3$, so we would have initially expected the smallest test error to be located at the matching k-value, $k=3$.
        However, the data generating process for the training data set labels \textit{and} the testing data set labels were injected with noise, so k-values greater than 3 have a ``de-noising'' effect allowing for better performance at around $k=7$.
        As the k-value increases beyond $k=7$, this smoothing effect is too strong, causing the gradual increase in test error as the k value increases.
        For $k<7$ and in particular $k<v=3$ we see the error growing because the model has small $k$ and is able to fit to the noise injected into the data generating process, corrupting the model.
        By averaging over 100 runs, the plot is relatively smooth allowing us to make generalisations about the relationship between k-value and mean test set error.
    \end{enumerate}
\newpage
\item[8.]
    \begin{enumerate}
        \item Producing a visualisation using Protocol B:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q8/q8}
            \caption{Protocol B}
            \label{fig:8}
            \end{figure}
        \item k-nearest neighbours is a non-parametric model, so when queried with a test point for prediction, the number of training points within a given fixed radius of that test point increases as the number of training points increases.
        If we consider this fixed radius `sufficiently close' to make an informed prediction, the number of points within this radius will increase with more training points.
        So as the number of training points increases, the optimal k would increase to include more of the training points within this same fixed radius.
        We can see in the plot above that we see this general trend as expected, however there is considerable noise in the trend and so a smoother plot would likely requires more runs.
    \end{enumerate}
\newpage
\section{PART III}
\subsection{Questions}
\item[9.] \begin{enumerate}
              \item let $K_{c}(\textbf{x},\textbf{z}) = c+ \sum_{i=1}^{n}x_{i}z_{i} = c + \textbf{x}^{T}\textbf{z}$\\
        Our function $K_{c}(\cdot, \cdot)$ is psd iff:\\
        $\forall \{ u_{i} \}_{1:m} \in \mathcal{R}, {\textbf{x}_{j}\}_{1:m} \in \mathcal{R}^{n}$:

        $\sum_{k,l} u_{k}u_{l}K_{c}(\textbf{x}_{u},\textbf{x}_{l}) \ge 0 $\\
        We expand our kernel function to give: \\
        \[\sum_{k,l} u_{k}u_{l}K_{c}(\textbf{x}_{u},\textbf{x}_{l}) =  \sum_{n,m} cu_{n} u_{m} + \sum_{n,m} \langle\textbf{x}_{n}, \textbf{x}_{m}\rangle u_{n} u_{m}\]

@@

        \textbf{Proposition}
        $K_{c}(\cdot, \cdot)$ is psd iff $c \ge0$ \\
        \textbf{Proof:}\\
        $(\implies)$\\

@@

    We note that as c becomes large, $\alpha \to 0$ and hence the predictions for $\hat{y}_{test}$ approach
    the constant function $f(x) = lc$\\

    Hence, adding c to our kernel acts like a regulariser in the limit.\\ 
\end{enumerate}

@@

    $K = I_{l} + \tilde{K}$, where $\tilde{K}_{ij} = K(x_i, x_j)$, $i \ne j$. We note here that as $\beta \to \infty$, $\tilde{K} \to 0$.\\

@@

    Let's assume that there exists a unique point x in our training set of minimal distance to our test point.\\
    Define $d_{i} := \|x_{test} - x_{i}\|^{2}$, and $d* := min_{i} d_{i}$\\
    $g(t) = \frac{y_{*} exp(-\beta d_{*})}{|\sum_{i} y_{i} exp(-\beta d_{i})|} + \sum_{i\ne *}\frac{y_{i} exp(-\beta d_{i})}{|\sum_{i} y_{i} exp(-\beta 
    d_{i})|} = \frac{y_{*}}{|y_{*} + \sum_{i \ne *} y_{i} exp(-\beta (d_{i} - d_{*}))|} + \sum_{i\ne *}\frac{y_{i} exp(-\beta (d_{i} - d_{*}))}{|y_{*} + \sum_{i \ne *} y_{i} exp(-\beta (d_{i} - d_{*}))|}$.\\

    Note for $d_{i} \ne d_{*}$, $exp(-\beta(d_{i} - d_{*})) \to 0$ as $\beta \to \infty$. Hence:\\
    $g(t) \to \frac{y_{*}}{|y_{*} + 0|} + \frac{0}{|y_{*} + 0|} = y_{*} = y_{i} : x_{i} = min_{i} \|x_{test} - x_{i}\|$ which is the required 1-NN predictor.\\

\newpage
\item[11.] \\
Given a whack-a-mole grid of size n, it is solvable if we can apply a sequence from our $n^{2}$ permissible moves that results in a blank screen. We first index all permissible moves by $T_{i,j}$, which has the effect of flipping the state of squares $(i,j),(i-1,j),(i+1,j),(i,j-1)(i,j+1)$, as long as each coordinate is still on the grid.\\
We may consider each grid as an element of the space of square matrices over $\mathcal{F}_{2}$, the finite field of order 2: under this construction, we see that the application of a move $T_{i,j}$ corresponds to the sum between the grid and a matrix $T_{ij}$, where $T_{ij}$, where the entries of $T_{ij}$ are zero everywhere except at $(i,j),(i-1,j),(i+1,j),(i,j-1)(i,j+1)$, if these values are present on the grid. For example, if n = 4, the transformation $T_{2,2}$ corresponds to the addition of the matrix $T_{22}$:\\
